---
title: 'STAT 632: Homework 6'
author: "Alle Mani Chandana"
date: "2024-04-14"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load the data set
library(readr)
county_votes <- read_csv("C:/Users/Mani chandana/Downloads/county_votes16.csv")

# Head of the data set
head(county_votes)

# summary of the data set
summary(county_votes)
```

### Exercise 1:

a)  

```{r}
# Simple logistic regression for trump_win as a repsonse variable and obama_pctvotes as a predictor variable.
glm1 <- glm(trump_win ~ obama_pctvotes, data = county_votes, family = binomial)

#summary of simple logistic regression
summary(glm1)
```

The equation for the estimated logistic regression is $$ \log\left(\frac{\hat{p}(x)}{1 - \hat{p}(x)}\right) = \hat{\beta}_0 + \hat{\beta}_1 *  \text{obama_pctvotes} $$

$$ \log\left(\frac{\hat{p}(x)}{1 - \hat{p}(x)}\right) = 1.493 - 0.017 * \text{obama_pctvotes} $$

Here, $\log\left(\frac{\hat{p}(x)}{1 - \hat{p}(x)}\right)$ is the estimated probability of trump winning, $\hat{\beta}_0$ & $\hat{\beta}_1$ are estimated coefficients.

b)  

```{r}
# Load the required library
library(ggplot2)

# Scatter plot of trump_win and obama_pctvotes and superimpose the fitted logistic curve for the estimated probability of Trump winning.
ggplot(county_votes, aes(obama_pctvotes, trump_win)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se=F) +
  xlab("Percent of votes cast for Obama in 2012") +
  ylab("Estimated Probability Trump Won") + theme_bw()
```

c)  

```{r}
# The estimated probability of Trump winning  in a county with obama pctvotes = 40, 50, and 60.
new_x <- data.frame(obama_pctvotes = c(40, 50, 60))
predict(glm1, newdata = new_x, type = "response")
```

d)  The estimated coefficient $\hat{\beta_1}$ indicates one unit increase in `obama_pctvotes` is associated with a change in the log-odds, or logit, by $\hat{\beta_1}$ = -0.371 (decreases the likelihood of `trump_win`).

### Exercise 2:

a)  

```{r}

# Multiple logistic regression model of trump_win as response varible and the remainig variables as predictor variables.
glm2 <- glm(trump_win ~ pct_pop65 + pct_black + pct_white + pct_hispanic + pct_asian + highschool + bachelors + income, data = county_votes, family = binomial)

# summary of mulitple logistic regression model.
summary(glm2)
```

b)  In the above multiple logistic regression model, there is only one insignificant variable, which is `pct_pop65`, since its p-value is greater than 0.05. Now, we update the multiple logistic regression model by removing this variable.

```{r}
# Updated multiple logistic regression model.
glm3 <- glm(trump_win ~ pct_black + pct_white + pct_hispanic + pct_asian + highschool + bachelors + income, data = county_votes, family = binomial)

# summary of updated mulitple logistic regression model.
summary(glm3)
```

c)  The negative sign of the estimated coefficients indicates that a unit increase in these variables (`pct_black`, `pct_hispanic`, `pct_asian`, `highschool`, `bachelors`) is associated with a change in the log-odds, or logit, by -0.0365, -0.0826, -0.1521, -0.0437, -0.1924, respectively.

The positive sign of the estimated coefficients indicates that a unit increase in these variables (`pct_white`, `income`) is associated with a change in the long-odds, or logit, by 0.0819, 0.0505, respectively.
